PhD Thesis

### 2016-12-20 pldi14/sec-practice.tex
Replace paper with chapter (with care)
Move appendix on evaluation into evaluation

Moreover, we can show that those axioms are consistent by
providing models implementing them. \pg{Resume}
%Additionally, we have two implementations of those axioms.
\pg{A couple more axioms need to be checked.}
%
% (s₂ ⊝ (ds ⊕ s₁)) ∘ ds &= s₂ ⊝ s₁ (?)
%
% Special case:
%	(ds ⊕ s) ⊝ s = ds (EDIT: only if ds is valid for s).
%
% Therorems:
%
%s ⊝ s = nil s
%(s₁ ⊝ s₂) ∘ (s₂ ⊝ s₃) = s₁ ⊝ s₃
\end{oldSec}

@@

For lack of space, details on benchmarking results and inputs
are available in the extended version of our paper (Appendix A).

>>>>

Moreover, we can show that those axioms are consistent by
providing models implementing them. \pg{Resume}
%Additionally, we have two implementations of those axioms.
\pg{A couple more axioms need to be checked.}
%
% (s₂ ⊝ (ds ⊕ s₁)) ∘ ds &= s₂ ⊝ s₁ (?)
%
% Special case:
%	(ds ⊕ s) ⊝ s = ds (EDIT: only if ds is valid for s).
%
% Therorems:
%
%s ⊝ s = nil s
%(s₁ ⊝ s₂) ∘ (s₂ ⊝ s₃) = s₁ ⊝ s₃
\end{oldSec}
\subsection{Benchmarking setup}
We run object language programs by generating corresponding Scala code.
To ensure rigorous
benchmarking~\citep{Georges07rigorousJavaPerformance}, we use the
Scalameter benchmarking library. To show that the performance
difference from the baseline is statistically significant, we
show 99\%-confidence intervals in graphs.

We verify \cref{eq:correctness} experimentally
by checking that the two sides of the equation always
evaluate to the same value.

We ran benchmarks on an 8-core Intel Core i7-2600 CPU running at
3.4 GHz, with 8GB of RAM, running Scientific Linux release 6.4.
While the benchmark code is single-threaded, the JVM offloads
garbage collection to other cores. We use the preinstalled
OpenJDK 1.7.0\_25 and Scala 2.10.2.

\paragraph{Input generation}

Inputs are randomly generated to resemble English words over all
webpages on the internet: The vocabulary size and the average
length of a webpage stay relatively the same, while the number of
webpages grows day by day. To generate a size-$n$ input of type
$(\HashMap{\Int}{\Bag*[\Int]})$, we generate $n$ random numbers
between 1 and 1000 and distribute them randomly in $n/1000$ bags.
Changes are randomly generated to resemble edits. A change has
50\% probability to delete a random existing number, and has 50\%
probability to insert a random number at a random location.

\paragraph{Experimental units}

Thanks to \cref{eq:correctness}, both recomputation
$\App{f}{\Apply*{\D a}{a}}$ and incremental computation
$\Apply{\App{\App{\Derive{f}}{a}}{\D a}}{\App{f}{a}}$ produce
the same result. Both programs are written in our object language.
To show that derivatives are faster, we compare
these two computations. To compare with recomputation, we measure the
\emph{aggregated} running time for running the derivative on the change
and for updating the original output with the result of the derivative.

@@

\pg{Warning: benchmark results are hardcoded here, copy-pasted
  from the spreadsheet. Edit with care.}

We present our results in \cref{fig:graph}. As expected, the
runtime of incremental computation is
\emph{essentially constant} in the size of the input, while the runtime
of recomputation is linear in the input size.
Hence, on our biggest inputs\pg{specify} incremental computation
is over $10^4$ times faster.

Derivative time is in fact slightly irregular for the first few inputs,
but this irregularity decreases slowly with increasing warmup
cycles. In the end, for derivatives we use $10^4$ warmup cycles.
With fewer warmup cycles, running time for derivatives decreases
significantly during execution, going from 2.6ms for $n = 1000$ to
0.2ms for $n = 512000$. Hence, we believe extended warmup is
appropriate, and the changes do not affect our general
conclusions. Considering confidence intervals, in our experiments the running
time for derivatives varies between 0.139ms and 0.378ms.

In our current implementation, the code of the generated
derivatives can become quite big. For the histogram example
(which is around 1KB of code), a pretty-print of its derivative
is around 40KB of code. The function application case in
\cref{fig:correctness:derive} can lead to a quadratic growth in
the worst case. More importantly, we realized that our home-grown
transformation system in some cases performs overly aggressive
inlining, especially for derivatives, even though this is not
required for incrementalization, and believe this explains a
significant part of the problem. Indeed, code blowup problems do
not currently appear in later experiments (see
\cref{ch:inc-more-programs}). \pg{Revise forward pointer.}

\paragraph{Summary}
Our results show that the incrementalized program runs
in essentially constant time and hence orders of magnitude faster than
the alternative of recomputation from scratch.
